### 2025-01-06 skripsi/chapter4.tex
Finish ch4, fix some problems with

% TODO: add figures and tables of generated data and some intermediate steps.
\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean_perturbed_solution.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_noise_levels}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_derivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:derivative}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_noise_levels}) A generated function with no noise (clean), low noise level (5\%), medium noise level (10\%), and high noise levels (50\%).\ (\subref{fig:antiderivative}) Antiderivative function with no noise.\ (\subref{fig:derivative}) Derivative function with no noise.}
\end{figure}
% TODO: interesting note that model trains better on function to coefficient task when the antiderivative is computed from the randomly generated derivative than if the derivative is computed from the randomly generated antiderivative. THIS NEEDS A REASON AND ANALYSIS SO MIGHT NOT BE NEEDED IN THIS WRITING. MAYBE A GOOD IDEA FOR A SEPARATE PIECE

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  import xarray as xr
  xr.open_zarr(
      "DATASET_ADDRESS",
      storage_options = {"token": "anon"},
  )
  \end{lstlisting}
  \caption{Example of using Xarray to access a publicly accessible Zarr dataset hosted on Google Cloud Storage Bucket.}\label{fig:access_remote_zarr}
\end{figure}

The dataset includes 62 variables with some variables also spanning 13 discrete vertical levels in the atmosphere. For our use, we will only be using the 2-meter temperature of the atmosphere. This is the air temperature 2 meters above the surface be it land, sea, or inland water. This variable is in units of kelvin. The values of this variable is obtained by interpolating between the model values at the surface and the lowest model vertical level. We also limit the time range to ... An example of the data that was retrieved can be seen in .
% TODO: talk about selecting the time here
% ECMWF description
% - Role
% - legitimacy
% - Weatherbench2
% data description
% - data assimilation, maybe reference a general description
% - variables
% - dimensions/coordinates (time, spatial) and resolution in each
% - climatology

@@

The second type of dataset is retrieved from an external source. The data source is first identified, in our case it is the WeatherBench 2 collection of datasets. The specific dataset from the collection of different dataset versions is first determined. The dataset is available in the \emph{Zarr} format. And then, we determine the filters for the portions of the dataset we are interested in. Once the dataset and filters are determined, the retrieval process starts by accessing the dataset using the \emph{Xarray} library \autocite{hoyer2017xarray,hoyerXarray2024}. The filters are applied as in .... % TODO: add figure for code of the filtering process and local save process
Finally, the filtered version is stored locally. The local dataset is also stored in the \emph{Zarr} format.

@@

% TODO: decide if hyperparameter selection is needed and explain why or why not. Maybe hyperparameter selection is done in a phase after the model evaluation with a decision between repeating model training with different hyperparameters or keeping the trained model.
Next, the selection of hyperparameters is done to optimize the model further. As previously mentioned, the hyperparameters are the regularization constant \(C \) and kernel functions including the kernel function parameters.

@@

The calculus operations for Fourier basis take advantage of the properties of the Fourier series. As discussed in \lccref{sec:fourier_discussion}, derivatives and integration become multiplication and division in the spectral space of Fourier series. Our implementation takes advantage of this as shown in \lccref{fig:basis_grad_impl}. Similar to how subtraction is just a modification of the addition operation, the integral or antiderivative operator is also implemented simply by changing the operation with the term multiplier from multiplication to division. % TODO: change the reference to specific equations for calculus in spectral domain

@@

The SpectralSVR class also provides a convenience function for testing the performance of the learned model. This function is shown in \lccref{fig:spectralsvr_test_impl}. The function arguments are the testing inputs, expected testing outputs, and a resolution for evaluation of function values. The first step is to convert the testing input into real representations and then predicting the output function. The predicted output are then compared to the testing outputs using many metrics including RMSE, MSE, MAE, R2, SMAPE, RRSE, and the number of Nan values present in the predicted output. The same comparison process is carried out again for the function values which are evaluated at the specified resolution. Finally, the function returns the metric values as a nested tuple for both the coefficients predictions and the function value predictions.

@@

\subsection{Scenario 1}
\noindent The first scenario results are presented in four parts. The first is the coefficient prediction metrics which is shown in \lccref{table:scenario_1_spectral_metrics}. The table shows the noise level and the performance metrics.
\begin{table}[H]
  \caption{Performance metrics of coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrrr}
    \toprule
    Noise level & MSE   & RMSE & MAE  & R2   & sMAPE \\
    \midrule
    5\%         & 0.86  & 0.93 & 0.57 & 0.96 & 0.31  \\
    10\%        & 2.06  & 1.43 & 0.93 & 0.91 & 0.46  \\
    50\%        & 15.59 & 3.95 & 3.04 & 0.50 & 1.04  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\Cref{table:scenario_1_spectral_metrics} shows that the model is able to generalize from the training data and learn the simple linear antiderivative operator. Focusing on the coefficient of determination (R2) scores, we can see that the model is off by 0.04 to the perfect score of 1.0 for the lowest noise level.  While increasing noise levels does degrade the performance, this is partly due to the target values also having been perturbed. This fact can be seen if we look at how predictions from the very same trained models on noisy inputs are compared to the noise-free versions of the antiderivative functions which is shown in \lccref{table:scenario_1_clean_spectral_metrics}. The metrics show that the model predictions are slightly closer to the unperturbed function coefficients compared to the perturbed versions. This can be seen as the influence of the independent synthetic measurement noise in the perturbed targets being removed from the testing metrics.

The higher noise levels shows that the model is still relatively accurate. Looking at the R2 score for the 50\% noise level, the model is more accurate than a baseline mean value model. The R2 score of 0.5 means that the model predictions has roughly half the error the baseline would produce.

\begin{table}[H]
  \caption{Performance metrics of coefficient prediction compared to unperturbed targets in scenario 1 by noise level.}\label{table:scenario_1_clean_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R2   & sMAPE \\
    \midrule
    5\%         & 0.79 & 0.89 & 0.51 & 0.96 & 0.28  \\
    10\%        & 1.80 & 1.34 & 0.80 & 0.92 & 0.40  \\
    50\%        & 9.30 & 3.05 & 2.23 & 0.62 & 0.90  \\
    \bottomrule
  \end{tabular}
\end{table}

The function values are also evaluated and the values of the metrics between the noisy targets and predictions are shown in \lccref{table:scenario_1_function_metrics}. While the relative metrics R2 and sMAPE stay close to the values we obtained for the coefficients themselves, since the metrics are now being computed in the physical domain, the scales have changed and because of that the values of the absolute metrics now reflect the change in scale. For reference, the scale of the target coefficient values for the 50\% noise level dataset in their real representation can range on average from -13.54 to 13.6 across the entire dataset. Therefore, an RMSE value of 3.94 for the predictions compared to the noisy targets mean that the error ratio is roughly 0.148 or just above 1 to 7. For the function values of the 50\% noise dataset the values range on average from -2.5 to 2.49. This was more difficult to compute since there is no direct way to compute the amplitude of the function from just the coefficients. Here we used the inverse transform and extracted the maximum and minimum values of each function from the discrete values. Since the RMSE in \lccref{table:scenario_1_function_metrics} for the 50\% noise dataset is 0.63, the error ratio comes to 0.126. One side note about \lccref{table:scenario_1_function_metrics} is that since the model instances are the same as the ones used in the coefficient evaluations, the hyperparameters are still the same including the kernel scaling factor sigma.
\begin{table}[H]
  \caption{Performance metrics of function value from evaluated coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_function_metrics}
  \centering
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R2   & sMAPE \\
    \midrule
    5\%         & 0.03 & 0.18 & 0.15 & 0.97 & 0.39  \\
    10\%        & 0.08 & 0.29 & 0.23 & 0.92 & 0.55  \\
    50\%        & 0.62 & 0.79 & 0.63 & 0.49 & 1.05  \\
    \bottomrule
  \end{tabular}
\end{table}

For the metrics between the predictions and the noise-free version of the target function values, the results parallel the outcomes of the coefficient predictions. The metrics show that the model performs slightly better with more pronounced differences for the higher noise levels. This echoes our previous statement on the effect of independent \enquote{measurement} noise in the target values on the evaluation metrics.
\begin{table}[H]
  \caption{Performance metrics of function value from evaluated coefficient prediction compared to unperturbed targets in scenario 1 by noise level.}\label{table:scenario_1_clean_function_metrics}
  \centering
  \begin{tabular}{lrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R2   & sMAPE \\
    \midrule
    5\%         & 0.03 & 0.18 & 0.14 & 0.97 & 0.38  \\
    10\%        & 0.07 & 0.27 & 0.21 & 0.93 & 0.53  \\
    50\%        & 0.37 & 0.61 & 0.49 & 0.62 & 0.94  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_exact_input.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_exact_input}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_exact_prediction.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_exact_prediction}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_exact_label}) The perturbed exact input function values from \lccref{eq:sine_derivative}.\ (\subref{fig:antiderivative_exact_prediction}) Prediction of antiderivative from the input function that was perturbed.}\label{fig:antiderivative_exact}
\end{figure}

@@

\begin{table}[H]
  \caption{Performance metrics of coefficient inverse prediction of derivative in scenario 1 by noise level.}\label{table:scenario_1_inv_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    Noise level & MSE      & RMSE   & MAE    & R2    & sMAPE \\
    \midrule
    5\%         & 1739.56  & 41.71  & 26.44  & 0.91  & 0.36  \\
    10\%        & 3485.10  & 59.03  & 42.19  & 0.76  & 0.52  \\
    50\%        & 49319.73 & 222.08 & 174.48 & -0.19 & 1.05  \\
    \bottomrule
  \end{tabular}
\end{table}

Next, we have the correlation image and p-matrix for the model of each noise level. The results are shown in \lccref{fig:scenario_1_interpretation}. Each row in the figure represent the results of a model trained on a different noise level. The correlation image for 5\% and 10\% noise show clear lines on the coefficient corresponding to the output coefficient that was sorted. We can also see that the negative wave number reflection also being sorted. This is because of the reflection that occurs with complex Fourier coefficients for real-valued functions. There are other faint vertical lines you are able to see for other input coefficients. These faint lines are not sorted like the corresponding coefficients we mentioned before. Meaning, while they don't affect the particular output coefficients that were sorted, they do show that there is information embedded in the kernel matrix for these input coefficients. However, if we look at the 50\% noise correlation image, even the corresponding coefficient does not show a clear line or gradient. The lines are more noisy compared to the other noise levels. But the kernel still manages to embed some information. The p-matrices show that The model itself is able to still learn some relationship between the input coefficients and the output coefficients. For the 5\% and 10\% noise levels, the p-matrices show very clearly the contributions of input coefficients to the output coefficients. However, looking at the p-matrix for 50\% noise level, the lower wave numbers show lower contribution of input coefficients to the output coefficients.
\begin{figure}[H]
  \begin{adjustwidth}{-0.1\linewidth}{-0.1\linewidth}
    \centering
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_ci_5.pgf}
      \end{adjustbox}
      \caption{Correlation image 5\% noise.}\label{fig:sc1_ci_5}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_pm_5.pgf}
      \end{adjustbox}
      \caption{The p-matrix for 5\% noise.}\label{fig:sc1_pm_5}
    \end{subfigure}
    % \\[\baselineskip]
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_ci_10.pgf}
      \end{adjustbox}
      \caption{Correlation image 10\% noise.}\label{fig:sc1_ci_10}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_pm_10.pgf}
      \end{adjustbox}
      \caption{The p-matrix for 10\% noise.}\label{fig:sc1_pm_10}
    \end{subfigure}
    \\[\baselineskip]
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_ci_50.pgf}
      \end{adjustbox}
      \caption{Correlation image 50\% noise.}\label{fig:sc1_ci_50}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_pm_50.pgf}
      \end{adjustbox}
      \caption{The p-matrix for 50\% noise.}\label{fig:sc1_pm_50}
    \end{subfigure}
  \end{adjustwidth}
  \caption{Correlation image (left column) and p-matrix (right column) for each model trained on a different noise level (row). The correlation image was sorted same order the values of the real component of wave number \(k=2\) were sorted in descending order.}\label{fig:scenario_1_interpretation}
\end{figure}

@@

\begin{table}[H]
  \caption{Performance metrics of coefficient prediction of next time step in scenario 2 by viscosity.}\label{table:scenario_2_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & R2   & sMAPE \\
    \midrule
    0.0      & 3.46e-04 & 1.86e-02 & 1.24e-02 & 0.79 & 0.65  \\
    0.01     & 3.56e-04 & 1.89e-02 & 1.27e-02 & 0.78 & 0.67  \\
    0.1      & 9.17e-04 & 3.03e-02 & 2.32e-02 & 0.46 & 1.08  \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean_perturbed_solution.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_noise_levels}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_derivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:derivative}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_noise_levels}) A generated function with no noise (clean), low noise level (5\%), medium noise level (10\%), and high noise levels (50\%).\ (\subref{fig:antiderivative}) Antiderivative function with no noise.\ (\subref{fig:derivative}) Derivative function with no noise.}
\end{figure}
% TODO: interesting note that model trains better on function to coefficient task when the antiderivative is computed from the randomly generated derivative than if the derivative is computed from the randomly generated antiderivative. THIS NEEDS A REASON AND ANALYSIS SO MIGHT NOT BE NEEDED IN THIS WRITING. MAYBE A GOOD IDEA FOR A SEPARATE PIECE

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  import xarray as xr
  ds = xr.open_zarr(
      "DATASET_ADDRESS",
      storage_options = {"token": "anon"},
  )
  \end{lstlisting}
  \caption{Example of using Xarray to access a publicly accessible Zarr dataset hosted on Google Cloud Storage Bucket.}\label{fig:access_remote_zarr}
\end{figure}

The specific dataset resolution we use is 64 by 32 representing longitude and latitude respectively. The ERA5 reanalysis dataset is hosted at the following URL\@: \url{gs://weatherbench2/datasets/era5/1959-2023_01_10-6h-64x32_equiangular_conservative.zarr}. The corresponding climatology is hosted at \url{gs://weatherbench2/datasets/era5-hourly-climatology/1990-2019_6h_64x32_equiangular_conservative.zarr}. The climatology is the average for hours 0, 6, 12, and 18 for each day of the year from 1990 until 2019. This average helps in defining the expected state of the weather at specific points in time based on the historical averages for that time of day and year.

For the reanalysis dataset, we limit the time range to the years 1995{-}2013 for training and 2020{-}2022 for testing. The specific time ranges for training are chosen to balance the diversity within the data and capacity of the machine the processing will be done on. The testing set time range was chosen to be as close to those shown on the WeatherBench 2 website. The filter for time and variables is shown in \lccref{fig:wb2_retrieval}. The reason for the different time ranges between training and testing is to reduce the risk of information leakage accross the training and testing sets. This ensures that the testing scores we measure will be for the most part from the generalization that model learned during training. The filter is an array the length of the number of samples. The array values are all the datetime values that the year ranges are true for. This results in an array of boolean values. Using this boolean filter we can index the data so that we get the filtered data. Then, the filtered data is saved locally so that we can access it in a more performant way. The encoding chunks may need to be removed in order to evade the Xarray library from mistakenly persisting the remote encoding after slicing data \autocite{ianZarrEncodingAttributes}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  years_idx = ds['time.year']
  data_idx = (years_idx > START_YEAR) & (years_idx < END_YEAR)
  da_orig = ds['2m_temperature'][data_idx]
  # save locally
  da_orig.encoding.pop('chunks',None)
  da_orig.to_zarr(DATA_FILENAME, mode='w')
  # open local version
  da = xr.open_zarr(DATA_FILENAME)
  \end{lstlisting}
  \caption{Example of filtering of Xarray dataset by year and variable. The data is saved and then loaded locally.}\label{fig:wb2_retrieval}
\end{figure}

The dataset includes 62 variables with some variables also spanning 13 discrete vertical levels in the atmosphere. For our use, we will only be using the 2-meter temperature of the atmosphere. This is the air temperature 2 meters above the surface be it land, sea, or inland water. This variable is in units of kelvin (K). Vertically, the values of this variable is obtained by interpolating between the model values at the surface and the lowest model vertical level. An example of the data that was retrieved can be seen in \lccref{fig:wb2_data}. The data shown is in the equiangular projection. This means that towards the poles, the shapes become more distorted. While this is not the most ideal way to represent the data, using the data in this way adds to the challenge of how the model can adapt to the use of suboptimal configurations in the data. To compute the anomaly, we can simply subtract the current state in \lccref{fig:wb2_train_data} by the climatology in \lccref{fig:wb2_clim_data} resulting in \lccref{fig:wb2_train_anomaly}.
\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/wb2_train_data.pgf}
    \end{adjustbox}
    \caption{}\label{fig:wb2_train_data}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/wb2_clim_data.pgf}
    \end{adjustbox}
    \caption{}\label{fig:wb2_clim_data}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/wb2_train_anomaly.pgf}
    \end{adjustbox}
    \caption{}\label{fig:wb2_train_anomaly}
  \end{subfigure}
  \caption{Data retrieved:(\subref{fig:wb2_train_data}) 2-meter temperature at January 1, 1995 6:00 UTC\@. (\subref{fig:wb2_train_anomaly}) Anomaly for 2-meter temperature at January 1, 1995 6:00 UTC\@. (\subref{fig:wb2_clim_data}) 2-meter temperature climatology for the first day of the year at hour 6 UTC\@.}\label{fig:wb2_data}
\end{figure}

After filtering, our training set has the dimensions of 27760 time steps, 64 longitude, and 32 latitude. The testing set on the other hand has 4384 time steps and the same values for the other two dimensions. For uses, such as a validation set during hyperparameter optimization, one can use a subset of the training set. The dimensions of the climatology are 4 for hour (0, 6, 12,\& 18), 366 for day of year, 64 for longitude, and 32 for latitude.

@@

The second type of dataset is retrieved from an external source. The data source is first identified, in our case it is the WeatherBench 2 collection of datasets. The specific dataset from the collection of different dataset versions is first determined. The dataset is available in the \emph{Zarr} format. And then, we determine the filters for the portions of the dataset we are interested in. Once the dataset and filters are determined, the retrieval process starts by accessing the dataset using the \emph{Xarray} library \autocite{hoyer2017xarray,hoyerXarray2024}. The filters are applied as in \lccref{fig:wb2_retrieval}. Finally, the filtered version is stored locally. The local dataset is also stored in the \emph{Zarr} format.

@@

For our study, we choose to not optimize the hyperparameters to focus on the broad performance with default hyperparameter values. The scaling parameter of the RBF kernel is computed from the training features. This is computed as the square root of the sum of feature variances. The regularization parameter used will be a default value of 1. Using the same values allow for easier comparisons for comparable results.

@@

The calculus operations for Fourier basis take advantage of the properties of the Fourier series. As discussed in \lccref{sec:fourier_discussion}, derivatives and integration become multiplication and division in the spectral space of Fourier series. Our implementation takes advantage of this as shown in \lccref{fig:basis_grad_impl}. Similar to how subtraction is just a modification of the addition operation, the integral or antiderivative operator is also implemented simply by changing the operation with the term multiplier from multiplication to division. % TODO: change the reference to specific equations for calculus in spectral domain in ch2

@@

The SpectralSVR class also provides a convenience function for testing the performance of the learned model. This function is shown in \lccref{fig:spectralsvr_test_impl}. The function arguments are the testing inputs, expected testing outputs, and a resolution for evaluation of function values. The first step is to convert the testing input into real representations and then predicting the output function. The predicted output are then compared to the testing outputs using many metrics including RMSE, MSE, MAE, R\textsuperscript{2}, SMAPE, RRSE, and the number of Nan values present in the predicted output. The same comparison process is carried out again for the function values which are evaluated at the specified resolution. Finally, the function returns the metric values as a nested tuple for both the coefficients predictions and the function value predictions.

@@

\subsection{Scenario 1}
\noindent The first scenario results are presented in four parts. The first is the coefficient prediction metrics which is shown in \lccref{table:scenario_1_spectral_metrics}. The table shows the noise level and the performance metrics.
\begin{table}[H]
  \caption{Performance metrics of coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrrr}
    \toprule
    Noise level & MSE   & RMSE & MAE  & R\textsuperscript{2}   & sMAPE \\
    \midrule
    5\%         & 0.86  & 0.93 & 0.57 & 0.96 & 0.31  \\
    10\%        & 2.06  & 1.43 & 0.93 & 0.91 & 0.46  \\
    50\%        & 15.59 & 3.95 & 3.04 & 0.50 & 1.04  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\Cref{table:scenario_1_spectral_metrics} shows that the model is able to generalize from the training data and learn the simple linear antiderivative operator. Focusing on the coefficient of determination (R\textsuperscript{2}) scores, we can see that the model is off by 0.04 to the perfect score of 1.0 for the lowest noise level.  While increasing noise levels does degrade the performance, this is partly due to the target values also having been perturbed. This fact can be seen if we look at how predictions from the very same trained models on noisy inputs are compared to the noise-free versions of the antiderivative functions which is shown in \lccref{table:scenario_1_clean_spectral_metrics}. The metrics show that the model predictions are slightly closer to the unperturbed function coefficients compared to the perturbed versions. This can be seen as the influence of the independent synthetic measurement noise in the perturbed targets being removed from the testing metrics.

The higher noise levels shows that the model is still relatively accurate. Looking at the R\textsuperscript{2} score for the 50\% noise level, the model is more accurate than a baseline mean value model. The R\textsuperscript{2} score of 0.5 means that the model predictions has roughly half the error the baseline would produce.

\begin{table}[H]
  \caption{Performance metrics of coefficient prediction compared to unperturbed targets in scenario 1 by noise level.}\label{table:scenario_1_clean_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R\textsuperscript{2}   & sMAPE \\
    \midrule
    5\%         & 0.79 & 0.89 & 0.51 & 0.96 & 0.28  \\
    10\%        & 1.80 & 1.34 & 0.80 & 0.92 & 0.40  \\
    50\%        & 9.30 & 3.05 & 2.23 & 0.62 & 0.90  \\
    \bottomrule
  \end{tabular}
\end{table}

The function values are also evaluated and the values of the metrics between the noisy targets and predictions are shown in \lccref{table:scenario_1_function_metrics}. While the relative metrics R\textsuperscript{2} and sMAPE stay close to the values we obtained for the coefficients themselves, since the metrics are now being computed in the physical domain, the scales have changed and because of that the values of the absolute metrics now reflect the change in scale. For reference, the scale of the target coefficient values for the 50\% noise level dataset in their real representation can range on average from -13.54 to 13.6 across the entire dataset. Therefore, an RMSE value of 3.94 for the predictions compared to the noisy targets mean that the error ratio is roughly 0.148 or just above 1 to 7. For the function values of the 50\% noise dataset the values range on average from -2.5 to 2.49. This was more difficult to compute since there is no direct way to compute the amplitude of the function from just the coefficients. Here we used the inverse transform and extracted the maximum and minimum values of each function from the discrete values. Since the RMSE in \lccref{table:scenario_1_function_metrics} for the 50\% noise dataset is 0.63, the error ratio comes to 0.126. One side note about \lccref{table:scenario_1_function_metrics} is that since the model instances are the same as the ones used in the coefficient evaluations, the hyperparameters are still the same including the kernel scaling factor sigma.
\begin{table}[H]
  \caption{Performance metrics of function value from evaluated coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_function_metrics}
  \centering
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R\textsuperscript{2}   & sMAPE \\
    \midrule
    5\%         & 0.03 & 0.18 & 0.15 & 0.97 & 0.39  \\
    10\%        & 0.08 & 0.29 & 0.23 & 0.92 & 0.55  \\
    50\%        & 0.62 & 0.79 & 0.63 & 0.49 & 1.05  \\
    \bottomrule
  \end{tabular}
\end{table}

For the metrics between the predictions and the noise-free version of the target function values, the results parallel the outcomes of the coefficient predictions. The metrics show that the model performs slightly better with more pronounced differences for the higher noise levels. This echoes our previous statement on the effect of independent \enquote{measurement} noise in the target values on the evaluation metrics.
\begin{table}[H]
  \caption{Performance metrics of function value from evaluated coefficient prediction compared to unperturbed targets in scenario 1 by noise level.}\label{table:scenario_1_clean_function_metrics}
  \centering
  \begin{tabular}{lrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R\textsuperscript{2}   & sMAPE \\
    \midrule
    5\%         & 0.03 & 0.18 & 0.14 & 0.97 & 0.38  \\
    10\%        & 0.07 & 0.27 & 0.21 & 0.93 & 0.53  \\
    50\%        & 0.37 & 0.61 & 0.49 & 0.62 & 0.94  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_exact_input.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_exact_input}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_exact_prediction.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_exact_prediction}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_exact_input}) The perturbed exact input function values from \lccref{eq:sine_derivative}.\ (\subref{fig:antiderivative_exact_prediction}) Prediction of antiderivative from the input function that was perturbed.}\label{fig:antiderivative_exact}
\end{figure}

@@

\begin{table}[H]
  \caption{Performance metrics of coefficient inverse prediction of derivative in scenario 1 by noise level.}\label{table:scenario_1_inv_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    Noise level & MSE      & RMSE   & MAE    & R\textsuperscript{2}    & sMAPE \\
    \midrule
    5\%         & 1739.56  & 41.71  & 26.44  & 0.91  & 0.36  \\
    10\%        & 3485.10  & 59.03  & 42.19  & 0.76  & 0.52  \\
    50\%        & 49319.73 & 222.08 & 174.48 & -0.19 & 1.05  \\
    \bottomrule
  \end{tabular}
\end{table}

Next, we have the correlation image and p-matrix for the model of each noise level. The results are shown in \lccref{fig:scenario_1_interpretation}. Each row in the figure represent the results of a model trained on a different noise level. The correlation image for 5\% and 10\% noise show clear lines on the coefficient corresponding to the output coefficient that was sorted. We can also see that the negative wave number reflection also being sorted. This is because of the reflection that occurs with complex Fourier coefficients for real-valued functions. There are other faint vertical lines you are able to see for other input coefficients. These faint lines are not sorted like the corresponding coefficients we mentioned before. Meaning, while they don't affect the particular output coefficients that were sorted, they do show that there is information embedded in the kernel matrix for these input coefficients. However, if we look at the 50\% noise correlation image, even the corresponding coefficient does not show a clear line or gradient. The lines are more noisy compared to the other noise levels. But the kernel still manages to embed some information. The p-matrices show that The model itself is able to still learn some relationship between the input coefficients and the output coefficients. For the 5\% and 10\% noise levels, the p-matrices show very clearly the contributions of input coefficients to the output coefficients. However, looking at the p-matrix for 50\% noise level, the lower wave numbers show lower contribution of input coefficients to the output coefficients.
\begin{figure}[H]
  \begin{adjustwidth}{-0.05\linewidth}{-0.05\linewidth}
    \centering
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_ci_5.pgf}
      \end{adjustbox}
      \caption{Correlation image 5\% noise.}\label{fig:sc1_ci_5}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_pm_5.pgf}
      \end{adjustbox}
      \caption{The p-matrix for 5\% noise.}\label{fig:sc1_pm_5}
    \end{subfigure}
    % \\[\baselineskip]
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_ci_10.pgf}
      \end{adjustbox}
      \caption{Correlation image 10\% noise.}\label{fig:sc1_ci_10}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_pm_10.pgf}
      \end{adjustbox}
      \caption{The p-matrix for 10\% noise.}\label{fig:sc1_pm_10}
    \end{subfigure}
    % \\[\baselineskip]
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_ci_50.pgf}
      \end{adjustbox}
      \caption{Correlation image 50\% noise.}\label{fig:sc1_ci_50}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/antiderivative_pm_50.pgf}
      \end{adjustbox}
      \caption{The p-matrix for 50\% noise.}\label{fig:sc1_pm_50}
    \end{subfigure}
  \end{adjustwidth}
  \caption{Correlation image (left column) and p-matrix (right column) for each model trained on a different noise level (row). The correlation image was sorted same order the values of the real component of wave number \(k=2\) were sorted in descending order.}\label{fig:scenario_1_interpretation}
\end{figure}

@@

\begin{table}[H]
  \caption{Performance metrics of coefficient prediction of next time step in scenario 2 by viscosity.}\label{table:scenario_2_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & R\textsuperscript{2}   & sMAPE \\
    \midrule
    0.0      & 3.46e-04 & 1.86e-02 & 1.24e-02 & 0.79 & 0.65  \\
    0.01     & 3.56e-04 & 1.89e-02 & 1.27e-02 & 0.78 & 0.67  \\
    0.1      & 9.17e-04 & 3.03e-02 & 2.32e-02 & 0.46 & 1.08  \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[H]
  \caption{Performance metrics of function values evaluated from coefficient prediction of next time step in scenario 2 by viscosity.}\label{table:scenario_2_function_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & R\textsuperscript{2}   & sMAPE \\
    \midrule
    0.0      & 5.58e-05 & 7.47e-03 & 5.89e-03 & 0.82 & 0.70  \\
    0.01     & 5.74e-05 & 7.58e-03 & 6.01e-03 & 0.81 & 0.71  \\
    0.1      & 1.58e-04 & 1.26e-02 & 9.98e-03 & 0.51 & 1.04  \\
    \bottomrule
  \end{tabular}
\end{table}
